---
title: "RMarkdown and Version Control with git"
author: "Lukas Schmid, Chân Lê"
date: "17 1 2020"
output: html_document
---

# __Introduction to R Markdown__

------

## __1. General__
RMarkdown can be used to save and execute codes, as well as generate high-quality reports to be shared with the audience.

A (newer version) cheatsheet of RMarkdown can be found here: https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf

## __2. Installation__

#### __a. Prerequisite__
Make sure that you have the __latest version of R and RStudio__. It was not officially mentioned on the RMarkdown official guide, but I did run into troubles trying to install RMarkdown without knowing that some conflicts could exist if you were not to get R in your computer updated.


#### __b. Install RMarkdown__
Next you can install RMarkdown package in R:

```{r eval = F}
# Install from CRAN
install.packages('rmarkdown')

# Or if you want to test the development version,
# install from GitHub
if (!requireNamespace("devtools"))
  install.packages('devtools')
devtools::install_github('rstudio/rmarkdown')
```


#### __c. LaTeX for PDF output__
If you want to generate PDF output, you will need to install LaTeX. For R Markdown users who have not installed LaTeX before, you can install TinyTeX (https://yihui.name/tinytex/):

```{r eval = F}
install.packages("tinytex")
tinytex::install_tinytex()  # install TinyTeX
```


## __3. Basics__

#### __a. Text formating__

The following code chunks present how text should look like when written in 
Markdown. To prevent the text from actually being formatted, we put it into
code chunks.

```{r eval = F}
Write normally for plain text 
```

Write normally for plain text

```{r eval = F}
> You can highlight some chunk of text or blockquotes with the greater-than sign
```

> You can highlight some chunk of text or blockquotes with the greater-than sign

```{r eval = F}
Italics using *single asterisks* or _single underscores_
```

Italics using *single asterisks* or _single underscores_

```{r eval = F}
Bold using **double asterisks** or __double underscore__
```

Bold using **double asterisks** or __double underscore__

```{r eval = F}
You can have different sizes of header with the number sign #

# Header 1   
## Header 2   
### Header 3   
#### Header 4   
##### Header 5   
###### Header 6

Alternatively, for H1, an underline-ish style:

Alt-H1

------
```

# Header 1   
## Header 2   
### Header 3   
#### Header 4   
##### Header 5   
###### Header 6

Alternatively, for H1, an underline-ish style:


Alt-H1

------

```{r eval = F}
Inline code blocks are highlighted using back-ticks, for example the package `ggplot2` 

```

Inline code blocks are highlighted using back-ticks, for example the package `ggplot2` 



```{r eval = F}
* You can create bullet points by setting an asterisk, just remember **the space between the asterisk and your text**:
  + Then add first sub-point with a plus sign
  + Another sub-point
    - Then a sub-sub-point using minus sign
```

* You can create bullet points by setting an asterisk, just remember **the space between the asterisk and your text**:
  + Then add first sub-point with a plus sign
  + Another sub-point
    - Then a sub-sub-point using minus sign
    
Bullet points need to have double line breaks before the former block of text. 
Continued lists have no double line breaks. Single line breaks will be ignored.

```{r eval = F}
1. You can create numbered list with, well, numbers.
2. Something something
   i) A sub-point (with three spaces at the beginning, to get the indentation)
   ii) Yet another sub-point
```

1. You can create numbered list with, well, numbers.
2. Something something
   i) A sub-point (with three spaces at the beginning, to get the indentation)
   ii) Yet another sub-point


### __b. Chunk code & Chunk options__

For all of this chapter and of course much more, see [https://bookdown.org/yihui/rmarkdown/r-code.html](chapter 2.6 in the
Definitive Guide to R Markdown.)

If you want to execute your code in RMarkdown, put them into code chunk brackets
(but don't normally put them as a comment - we just did this to demonstrate how
it should look). You can get the brackes by  ___pressing Ctrl + ALt + I__ ).

```{r eval = F}
  #```{r}
  test <- rnorm(20, mean = 0, sd = 3)
  mean(test)
  sd(test)
  # ```
```

Chunks have chunk options that change their behaviour when knitted. By default,
code in all chunks will be run by R, the code will be displayed in the final
Markdown-document and the outputs (if any) will be displayed. There are some
useful options to change this behaviour (standard behaviour is mentioned):

- `eval = TRUE`: Run code in chunk
- `echo = TRUE`: display the source code in the document
- `results = markup`: how to display the results of a code chunk
- `include = TRUE`: Include anything from the code at all (`include = FALSE` is equal to `echo = FALSE, results = 'hide', warning = FALSE, message = FALSE`)

Additionally, the first option in a code chunk can be the label of that chunk.
Labels are useful when you have lots of them and try to debug your document; 
they will also be used for filenames in plots.

And finally, you can also run code inline. So, if I want to report the mean of 
some variable directly in the text, I can write `#r mean(cars$speed)` (without
the hashtag) to get the result, `r mean(cars$speed)`.

### visualisation formatting


### table and reporting formatting
### document types and presentations)


# version control with git
## general principle
## GUI "GitKraken"
## usage from within RStudio
this is Lukas' Text


##hello from Chan

Lukas Schmid  vor 18 Minuten