---
title: "RMarkdown and Version Control with git"
author: "Lukas Schmid, Chân Lê"
date: "17 1 2020"
output: html_document
---

# __Introduction to R Markdown__

------

### __1. General__
RMarkdown can be used to save and execute codes, as well as generate high-quality reports to be shared with the audience.


### __2. Installation__

#### __a. Prerequisite__
Make sure that you have the __latest version of R and RStudio__. It was not officially mentioned on the RMarkdown official guide, but I did run into troubles trying to install RMarkdown without knowing that some conflicts could exist if you were not to get R in your computer updated.


#### __b. Install RMarkdown__
Next you can install RMarkdown package in R:

```{r eval = F}
# Install from CRAN
install.packages('rmarkdown')

# Or if you want to test the development version,
# install from GitHub
if (!requireNamespace("devtools"))
  install.packages('devtools')
devtools::install_github('rstudio/rmarkdown')
```


#### __c. LaTeX for PDF output__
If you want to generate PDF output, you will need to install LaTeX. For R Markdown users who have not installed LaTeX before, you can install TinyTeX (https://yihui.name/tinytex/):

```{r eval = F}
install.packages("tinytex")
tinytex::install_tinytex()  # install TinyTeX
```


### __3. Basics__

#### __a. Text formating__

```{r eval = F}
Write normally for plain text 
```

Write normally for plain text

```{r eval = F}
> You can highlight some chunk of text or blockquotes with the greater-than sign
```

> You can highlight some chunk of text or blockquotes with the greater-than sign

```{r eval = F}
Italics using *single asterisks* or _single underscores_
```

Italics using *single asterisks* or _single underscores_

```{r eval = F}
Bold using **double asterisks** or __double underscore__
```

Bold using **double asterisks** or __double underscore__

```{r eval = F}
You can have different sizes of header with the number sign #

# Header 1   
## Header 2   
### Header 3   
#### Header 4   
##### Header 5   
###### Header 6

Alternatively, for H1 and H2, an underline-ish style:

Alt-H1

======

Alt-H2

------
```

# Header 1   
## Header 2   
### Header 3   
#### Header 4   
##### Header 5   
###### Header 6

Alternatively, for H1 and H2, an underline-ish style:

### Alt-H2

------

```{r eval = F}
Inline code blocks are highlighted using back-ticks, for example the package `ggplot2` 

```

Inline code blocks are highlighted using back-ticks, for example the package `ggplot2` 



```{r eval = F}
* You can create bullet points by setting an asterisk, just remember **the space between the asterisk and your text**:
  + Then add first sub-point with a plus sign
  + Another sub-point
    - Then a sub-sub-point using minus sign
```

* You can create bullet points by setting an asterisk, just remember **the space between the asterisk and your text**:
  + Then add first sub-point with a plus sign
  + Another sub-point
    - Then a sub-sub-point using minus sign

```{r eval = F}
1. You can create numbered list with, well, numbers.
2. Something something
   i) A sub-point (with three spaces at the beginning, to get the indentation)
   ii) Yet another sub-point
```

1. You can create numbered list with, well, numbers.
2. Something something
   i) A sub-point (with three spaces at the beginning, to get the indentation)
   ii) Yet another sub-point


### __b. Chunk code & Chunk options__

If you want to execute your code in RMarkdown, put them into the following code chunk brackets ( __or press Ctrl + ALt + I__ ). For example:

```{r}
  test <- rnorm(20, mean = 0, sd = 3)
  mean(test)
  sd(test)
```









A (newer version)cheatsheet of can be found here: https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf



### text formatting

### chunk options for code
### visualisation formatting
### table and reporting formatting
### document types and presentations)


# version control with git
## general principle
## GUI "GitKraken"
## usage from within RStudio
this is Lukas' Text


##hello from Chan

Lukas Schmid  vor 18 Minuten