---
title: "RMarkdown and Version Control with git"
author: "Lukas Schmid, Chân Lê"
date: "17 1 2020"
output:
  html_document:
    toc: true
---

# __Introduction to R Markdown__

------

## __1. General__
RMarkdown can be used to save and execute codes, as well as generate high-quality reports to be shared with the audience.


## __2. Installation__

### __a. Prerequisite__
Make sure that you have the __latest version of R and RStudio__. It was not officially mentioned on the RMarkdown official guide, but I did run into troubles trying to install RMarkdown without knowing that some conflicts could exist if you were not to get R in your computer updated.


### __b. Install RMarkdown__
Next you can install RMarkdown package in R:

```{r eval = F}
# Install from CRAN
install.packages('rmarkdown')

# Or if you want to test the development version,
# install from GitHub
if (!requireNamespace("devtools"))
  install.packages('devtools')
devtools::install_github('rstudio/rmarkdown')
```


### __c. LaTeX for PDF output__
If you want to generate PDF output, you will need to install LaTeX. For R Markdown users who have not installed LaTeX before, you can install TinyTeX (https://yihui.name/tinytex/):

```{r eval = F}
install.packages("tinytex")
tinytex::install_tinytex()  # install TinyTeX
```


## __3. Basics__

### __a. Text formating__

Write normally for plain text

> Highlight some chunk of text or blockquotes with the greater-than sign

Write in italics using *single asterisks* or _single underscores_

Wrie in bold using **double asterisks** or __double underscore__

You can have different sizes of header with the number sign #:

# Header 1   
## Header 2   
### Header 3   
#### Header 4   
##### Header 5   
###### Header 6

Alternatively, you can add a horinzontal line to separate sections of your report:

------

Inline code blocks are highlighted using back-ticks, for example the package `ggplot2`


* You can create bullet points by setting an asterisk, just remember **the space between the asterisk and your text**:
  + Then add first sub-point with a plus sign
  + Another sub-point
    - Then a sub-sub-point using minus sign
    

1. You can create numbered list with, well, numbers.
2. Something something
   i) A sub-point (with three spaces at the beginning, to get the indentation)
   ii) Yet another sub-point
   
And finally, you can also run code inline. So, if I want to report the mean of 
some variable directly in the text, I can write `#r mean(cars$speed)` (without
the hashtag) to get the result, `r mean(cars$speed)`.

### __b. Chunk code & Chunk options__

If you want to execute your code in RMarkdown, put them into the following code chunk brackets ( __or press Ctrl + ALt + I__ ). The codes are displayed in a gray box, while the outputs are in a white box. For example:

```{r}
  library(tidyr)
  library(ggplot2)
  paste("Welcome","to","RMarkDown!")
  data <- iris
  data.plot <- data %>% pivot_longer(-Species, names_to = "Measurement", values_to = "Value")
  ggplot(data.plot, aes(group = Measurement, x = Measurement, y = Value)) + 
    geom_boxplot() +
    facet_grid(. ~ Species) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  lm.model <- lm(Petal.Length ~ Sepal.Length + Sepal.Width + Petal.Width, data = data)
  summary(lm.model)
```

Try not to crowd your code chunk like that. Instead of that, divide your work into smaller steps so you can comment on them easily.

On this second attempt, we you should you some essential code chunk options to make your report more clear and concise.

* First, I want to call some needed packages. If you want to __run these codes but not include them in the report__, type in chunk option `include = FALSE`

```{r include = FALSE}
library(tidyr)
library(ggplot2)
```

> The code ran but nothing appeared

* If you only want to show some __codes without running them__, type in chunk option `eval = FALSE`

```{r eval = FALSE}
  paste("Welcome","to","RMarkDown!")
```

* On the other hand, if you want to show some __results without the codes__, type in chunk option `echo = FALSE`

```{r echo = FALSE}
  paste("Welcome","to","RMarkDown!")
```

__Summary:__ Chunks have chunk options that change their behaviour when knitted. By default,
code in all chunks will be run by R, the code will be displayed in the final
Markdown-document and the outputs (if any) will be displayed. There are some
useful options to change this behaviour (standard behaviour is mentioned):

- `eval = TRUE`: Run code in chunk
- `echo = TRUE`: display the source code in the document
- `results = markup`: how to display the results of a code chunk
- `include = TRUE`: Include anything from the code at all (`include = FALSE` is equal to `echo = FALSE, results = 'hide', warning = FALSE, message = FALSE`)

Additionally, the first option in a code chunk can be the label of that chunk.
Labels are useful when you have lots of them and try to debug your document; 
they will also be used for filenames in plots.

```{r Demonstration_of_Label}
  paste("Welcome","to","RMarkDown!")
```

* Now I can load the data and view it.

```{r}
  data <- iris
  head(data)
```

* The next step is to transform it into long-format for the plotting:

```{r}
  data.plot <- data %>% pivot_longer(-Species, names_to = "Measurement", values_to = "Value")
  head(data.plot)
```

* Now I plot it. Multiple bar plots showing the distribution of different measurements of the iris flowers, facetting among different species. To align this plot in the middle, add chunk option `align = center` (others options including `left` and `right`). To resize the plot, add chunk option `fig.width` and `fig.height` with chosen values (in inches). Chunk option `fig.cap` is used to insert a string caption below the graph.

```{r fig.align="center", fig.height= 5, fig.width= 7, fig.cap = "Figure 1: possible caption for the graph"}
  ggplot(data.plot, aes(group = Measurement, x = Measurement, y = Value)) + 
    geom_boxplot() +
    facet_grid(. ~ Species) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### __c. Tables in RMarkdown using `knitr` and `kableExtra`__

As a default, you can always show a data set by just calling it directly. The dataset is shorten for the purpose of demonstration.

```{r}
  data <- head(data, 10)
  data
```

There is another alternative to displaying a data set, or broadly speaking, a table. With `knitr` and `kableExtra`

```{r}
  #install.packages("knitr")
  #install.packages("kableExtra")
  library(knitr)
  library(kableExtra)
  head(data) %>% 
    kable() %>% 
    kable_styling()
  
```

There are numerous features and layers that you can add to this table (same logic as in ggplot2). We would like to
only show one example that we think is relevant for the report: You can hightlight columns or rows in your table with `column_spec()` and `row_spec()`. For example:

```{r}
  head(data) %>% 
    kable() %>% 
    kable_styling() %>% 
    column_spec(1:2, #Specify the columns you wish to highlight
                bold = T, #Whether text turns bold
                color = "white", #Text color
                background = "blue") #Background color
```

```{r}
  head(data) %>% 
    kable() %>% 
    kable_styling() %>% 
    row_spec(3:5, #Specify the rows you wish to highlight
                bold = T, #Whether text turns bold
                color = "red", #Text color
                background = "green") #Background color
```


Other advanced formatting of tables (e.g. their alignement, stretching over multiple
pages...) in PDFs can be more cumbersome. Here, we only reference to
[https://bookdown.org/yihui/rmarkdown/r-code.html#tables](chapter
2.6.2 in the Markdown Guide). Vignettes on how to use kableExtra can be found on
the [CRAN-page](https://cran.r-project.org/web/packages/kableExtra/index.html).

When it comes to reporting results, plain `knitR` is not enough. There are
multiple packages available to do this ([see some examples
here](https://rmarkdown.rstudio.com/lesson-7.html)); I have used `stargazer` in
the past, which is especially useful in PFDs. It generates HTML-Code (or LaTeX-
code) directly from R-code. Due to this, we need a workaround: we always want
the result of stargazer to be displayed as is. This way, the results (now in)
HTML) will be read by the browser to be displayed as a table. An example helps:

```{r}
# This will only give us some HTML-formatted text:
stargazer::stargazer(lm(speed ~ dist, cars), type ="html")
```

```{r results='asis'}
# We now have the chunk-option results='asis':
stargazer::stargazer(lm(speed ~ dist, cars), type ="html")
```

## __4. YAML__

Every markdown has a header. The format of that header is called YAML (which
is short for the wonderful recursive acronym "YAML ain't markup language"). Most
important for us are the output options (
[chapter 2.4](https://bookdown.org/yihui/rmarkdown/output-formats.html) in the
definitive markdown guide). RMarkdown relies on Pandoc to convert between
different formats.

### __a. Output Types__

We can knit the Markdown-document into a wide range of files and, for many file
endings, there are multiple packages available. All of them are 'normal' R-
packages; our current one is `html_document`, so we can call help for it by
using `?html_document`. Among the most used probably are PDF and HTML, though
Powerpoint and Word are also possible.

You can not only write documents with RMarkdown. Another (very realistic) use 
case are slides. These work exactly like all other documents (can be knitted to
a variety of formats), with the addition of some elemtns necessary for slides.
This way, it is possible to use R and report results directly in your R-
environment.

### __a. Output Options__

It is not in the scope of this introduction to go into details here. However,
there is a structure underlying the whole thing. A plain YAML-header looks like 
this:

```{r eval=F}
---
title: "RMarkdown and Version Control with git"
author: "Lukas Schmid, Chân Lê"
date: "17 1 2020"
output: html_document
---
```

A quick look at the help tells us that we can insert an toc (Table of Contents)
into the document, and we cleary want that! So we can change the header 
accordingly:

```{r eval=F}
---
title: "RMarkdown and Version Control with git"
author: "Lukas Schmid, Chân Lê"
date: "17 1 2020"
output:
  html_document:
    toc: true
---
```

Depending on the package used and the output format, there are much more 
possibilities. I personally like the floating toc (`toc_float`).

## __5. Remarks & Literature__

* A (newer version)cheatsheet of can be found here: https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf

-------

# __Version control with git__

## __1. General principle__

Git is a version control software for line-based data (that means, you can
version control things like .txt, .csv, LaTeX, R, and other code, but not 
formats like .PDF, .docx, .xlsx, ar images). It is useful because it is (comparatively) simple and not storage-intensive.

You can use it locally, but you can also connect it to specialised websites 
(like  [GitHub](https://github.com/) or [BitBucket](https://bitbucket.org/)) - 
this ultimately means that you have an opportunity to back-up all of your code,
for no money, on the internet (and I think it is relatively secure, so your
code will not be stolen). GitHub also has a good 
[education plan](https://education.github.com/pack) which gives you
access to some cool programs.

## __2. How it works__

Many examples on the web use git from your command line. This is probably a
skill worth to learn at some point; however, you can do without - and as one 
website out it, "no one is handing out nerd badges", so using the command line 
because it looks cool is probably stupid (though we all do it, of course). So
here come only the basic basics.

### __a. Creating a repsository__

Git relies on so-called repositories. For us, they are simply folders that 
contain an additional `.git` folder and sometimes a `.gitignore` file, and of 
course everything you cant to version-control.

You can create a repository locally on your computer. If you want to use that
repository online, you can later push it to an online service. If you want to
use it online from the start, better create the repository online and then 
'clone' it to your machine.

### __b. Applying changes__

Whenever something changes in your repository, you have to first **add** the changes to your index. Because the index is the staging area (german: Zwischenschritt), this is also called **staging**. After changes are added, they
are not yet safed! To actually apply changes, you have to **commit** them. When
committing, you have to specify a commit message that usually contains info 
about what you have done in that commit. 

If you use an online repo, you have to **push** those changes (to the server) so
they not only exist on your local machine.

### __c. Workflow__

Of course, if you work online, you also have to check for changes on the server
(though this is less of an issue if you work alone on only one machine). This is
called to **pull**. As a result, we can summarise the following ideal workflow:

1. Pull to sync your local repo with the online repo.
2. Work.
3. Stage changes (whenever it seems reasonable; these can be very small steps!
The more and smaller steps you stage and commit, the higher is the change that
you can return to a well working version of your code.)
4. Commit the changes.
5. Push the changes.
6. Repeat 2 to 6 until you're done.

### __d. Branches and merges__

When you pull and there are changes in the online repo, these changes will
automatically be applied to your code. In some cases (e.g. the same line has
been changed in the online and the local repo, but in different ways), you have
to manually resolve these merge conflicts. This usually means to decide which
line should be used.

When more than one person works on a code, it is advisable to branch. This means
that effects in one branch won't affect any other branch. However, due to time
and ressource constraints, we won't cover this topic here.

## __3. Appendix: glossary__

- working directory: contains all your files
- repo: short for repository
- to clone a repo: copy an existing online-repo to your local machine
- HEAD: the last version of your repo (or the branch you are currently in); not
the same as the working directory
- index: the staging area where you 'park' changes until you commit them
- to stage or add change: to apply changes to the index
- to commit changes: to add changes from the index to the HEAD, thus making them
visible 
- to push changes: to add changes from your local repo to the online repo
- to pull changes: to add changes from your online repo to the local repo


